---
title: "sandbox-data"
author: "JJayes"
date: "19/05/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(glue)

```

## Purpose

To create some summary data from the data LR sent me - I'll use this summary data in the presentation on data viz.

```{r}
df <- read_rds("data/df_clean.rds")

df <- df %>% 
  select(`Parish name`, 
         Year_baptized, Month_baptized, Day_baptized, 
         Christian_name, Surname_b,
         Year_of_birth, Month_of_birth, Day_of_birth, 
         Race, 
         Female_age, Female_surname, Female_name, Female_country, Female_profession,
         Male_age, Male_surname, Male_name, Male_country, Male_Profession)

df <- df %>% 
  janitor::clean_names()

df <- df %>% 
  mutate(parish_address = case_when(
    
    str_detect(parish_name, "Christ Church") ~ "16 Summerley Rd, Kenilworth, Cape Town, 7708, South Africa",
    str_detect(parish_name, "St Mark") ~ "30 Bamford Ave, Athlone, Cape Town, 7760, South Africa",
    str_detect(parish_name, "St Peter") ~ "Victoria Rd, Mowbray, Cape Town, 7700, South Africa",
    str_detect(parish_name, "St James") ~ "St James Rd, Sea Point, Cape Town, 8005, South Africa",
    str_detect(parish_name, "St Cyprian") ~ "2B Papu St, Langa, Cape Town, 7456, South Africa",
    str_detect(parish_name, "Maitland") ~ "3 Fontana St, Brooklyn, Cape Town, 7405, South Africa",
    str_detect(parish_name, "St Margaret") ~ "25 Jansen St, Cape Town, 7500, South Africa",
    TRUE ~ " "
  ))

df %>% 
  count(male_country, sort = T) 

df %>% 
  count(female_country, sort = T) 

df %>% 
  count(male_profession, sort = T) 

df %>% 
  count(female_profession, sort = T)

df %>% distinct(parish_address)

df %>% distinct(parish_name)

df <- df %>% 
  mutate(parish_name = case_when(
    
    str_detect(parish_name, "Christ Church") ~ "Christ Church, Kenilworth",
    str_detect(parish_name, "St Mark") ~ "St Mark's Atholone",
    str_detect(parish_name, "St Peter") ~ "St Peter's, Mowbray",
    str_detect(parish_name, "St James") ~ "St James the Great, Sea Point",
    str_detect(parish_name, "St Cyprian") ~ "St Cyprian's, Langa",
    str_detect(parish_name, "Maitland") ~ "St Anne's, Maitland",
    str_detect(parish_name, "St Margaret") ~ "St Margaret's, Parow",
    TRUE ~ " "
  ))

df %>% distinct(parish_name)

df %>% 
  drop_na()

df %>% count(race, surname_b, sort = T) %>% 
  group_by(race) %>% 
  slice_max(n, n = 5, with_ties = F) %>% 
  ungroup() %>% view()

# write_rds(df, "data/baptism_data.rds")

```

# Using Baptism data

## Reading in data

```{r}
df <- read_rds("data/baptism_data.rds")

skimr::skim(df)
```

## Counts

```{r}
df %>% 
  count(parish_name, sort = T)

df %>% 
  count(year_of_birth, sort = T)

df %>% 
  count(year_baptized, sort = T)

```

## Lines

```{r}

df %>% 
  group_by(year_baptized, parish_name) %>% 
  mutate(number_of_baptisms = n()) %>% 
  ungroup() %>% 
  ggplot(aes(x = year_baptized, y = number_of_baptisms, colour = parish_name)) +
  geom_line() +
  scale_color_brewer(palette = "Dark2") +
  labs(x = NULL,
       y = "Number of baptisms per year")

```

```{r}
df %>% 
  group_by(year_baptized) %>% 
  mutate(number_of_baptisms = n()) %>% 
  ungroup() %>% 
  ggplot(aes(x = year_baptized, y = number_of_baptisms)) +
  geom_line() +
  facet_wrap(~ parish_name)
```


## Density of baptisms by day of the week

```{r}
df %>% 
  count()

```



### Map

How to use tidy geocoder

```{r}
library(tidygeocoder)

# locations <- df %>%
#   select(parish_address) %>%
#   distinct() %>%
#   tidygeocoder::geocode(address = parish_address,
#           method = "osm")
# 
# write_rds(locations, "data/locations.rds")

locations <- read_rds("data/locations.rds")

df <- df %>% 
  inner_join(locations, by = "parish_address")

locations %>% 
  ggplot(aes(long, lat, label = parish_address)) +
  geom_point() +
  geom_text()
```

### Basic map with topography

```{r}
library(ggmap)
locations %>% 
  skimr::skim()

# create boudning box for map image
b_box <- c(left = 18.35, bottom = -34.02, right = 18.65, top = -33.85)

# create map
CT_map <- get_map(b_box, zoom = 11)
```

### Nice!

```{r}
ggmap(CT_map) +
    geom_point(aes(long, lat), 
               data = locations) +
  geom_text(aes(x = long, y = lat,
                label = parish_name), 
               data = df %>% distinct(parish_name, long, lat))

```

### What can we fix?

```{r}

p <- ggmap(CT_map) +
    geom_point(aes(long, lat), 
               data = locations,
               size = 3,
               show.legend = F) +
  geom_text(aes(x = long, y = lat,
                label = parish_name), 
            vjust = 1.1,
               data = df %>% distinct(parish_name, long, lat) %>% 
              mutate(parish_name = str_remove(parish_name, "Cape Town,"),
                     parish_name = str_replace(parish_name, ",", "\n"))) +
  theme_void() +
  labs(title = "Where are our churches located?")

# write_rds(p, "data/map_of_churches.rds")
```

### Can we add more meaningful information to the plot?

- How many observations are there per church?

```{r}

df %>% 
  group_by(parish_name) %>% 
  summarise(n_obs = n()) %>% 
  arrange(desc(n_obs))

df %>% 
  group_by(parish_name) %>% 
  summarise(n_obs = n()) %>% 
  mutate(parish_name = fct_reorder(parish_name, n_obs)) %>% 
  ggplot(aes(x = n_obs, y = parish_name, fill = parish_name)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(x = n_obs, y = parish_name, label = n_obs), hjust = -.2) +
  scale_fill_brewer(palette = "Dark2") +
  expand_limits(x = 850) +
  labs(x = "Number of baptisms",
       y = NULL) +
  theme(panel.grid = element_blank(),
        panel.border = element_blank())
  

```


```{r}
df_map <- df %>% 
  group_by(parish_name) %>% 
  mutate(n_obs = n()) %>% 
  ungroup() %>% 
  distinct(parish_name, lat, long, n_obs)
```

### Reuse

We can reuse what we created earlier and just change the size argument to n_obs.

```{r}
ggmap(CT_map) +
    geom_point(aes(long, lat, size = n_obs), 
               data = df_map,
               show.legend = F) +
  geom_text(aes(x = long, y = lat,
                label = parish_name), 
            vjust = 1.1,
               data = df %>% distinct(parish_name, long, lat) %>% 
              mutate(parish_name = str_remove(parish_name, "Cape Town,"),
                     parish_name = str_replace(parish_name, ",", "\n"))) +
  theme_void() +
  labs(title = "Where are our churches located?")  
  
```


Leaflet map

```{r}

df <- df %>% 
  # this popup will contain the parsh name in bold (<b> ... </b> makes bold) 
  # and the address in the next line (with a <br/> break)
  mutate(popup = glue("<b>{parish_name}</b><br/>{parish_address}"))

df %>% 
  # makes a leaflet map
  leaflet() %>% 
  # adds a background
  addTiles() %>% 
  # sets the centre of the map with the mean values of lat and long
  setView(mean(df$long), mean(df$lat), zoom = 11) %>% 
  # here we add a marker with the popup specified above, inclduing the name of the parish and its address.
  addMarkers(unique(df$long), unique(df$lat), 
             popup = unique(df$popup))

```



### Line chart

```{r}
df %>% 
  group_by(parish_name, year_baptized) %>% 
  mutate(n_obs = n()) %>% 
  ungroup() %>% 
  distinct(parish_name, lat, long, n_obs, year_baptized) %>% 
  ggplot(aes(year_baptized, n_obs, colour = parish_name)) +
  geom_line() +
  geom_point()
```

### Small multiples

Key ideas = facet_wrap and remove legend with theme(legend.position = "none")

```{r}
df %>% 
  mutate(parish_name = str_remove(parish_name, "Cape Town,")) %>% 
  group_by(parish_name, year_baptized) %>% 
  mutate(n_obs = n()) %>% 
  ungroup() %>% 
  distinct(parish_name, lat, long, n_obs, year_baptized) %>% 
  ggplot(aes(year_baptized, n_obs, colour = parish_name)) +
  geom_line() +
  geom_point() +
  facet_wrap(~ parish_name) +
  theme(legend.position = "none")

```

### Geom jitter for map over time - a fun one for LinkedIn!

```{r}
df %>% 
  mutate(parish_name = str_remove(parish_name, "Cape Town,")) %>% 
  group_by(parish_name) %>% 
  mutate(n_obs = n()) %>% 
  ungroup() %>% 
  select(parish_name, lat, long, n_obs) %>% 
  ggplot(aes(long, lat)) +
  geom_jitter(width = .01,
              height = .01)

df %>% 
  mutate(parish_name = str_remove(parish_name, "Cape Town,")) %>% 
  group_by(parish_name, year_baptized) %>% 
  mutate(n_obs = n()) %>% 
  ungroup() %>% 
  distinct(parish_name, lat, long, n_obs, year_baptized)


```


### lollipop chart with average age of each spouse by race and by parish

```{r}

df %>% 
  filter(!is.na(female_age),
         !is.na(male_age)) %>% 
  group_by(race) %>% 
  summarise(mean_female_age = mean(female_age),
            mean_male_age = mean(male_age)) %>% 
  ggplot() +
  geom_point(aes(mean_female_age, race), colour = "blue") +
  geom_point(aes(mean_male_age, race), colour = "red") +
  labs(x = "Mean age of spouse",
       y = NULL,
       title = "Average age gap between husband and wife by race")

```

### ggridges plot

```{r}

df %>% 
  filter(!is.na(female_age),
         !is.na(male_age)) %>% 
  group_by(race) %>% 
  summarise(female = mean(female_age),
            male = mean(male_age),
            n = n()) %>% 
  pivot_longer(c(male, female), names_to = "gender") %>% 
  mutate(race = glue::glue("{race} - {n}"),
         race = fct_reorder(race, n)) %>% 
  ggplot(aes(value, race, colour = gender)) +
  geom_point() 

```


```{r}

df %>% 
  filter(!is.na(female_age),
         !is.na(male_age)) %>% 
  ggplot(aes(female_age, fill = race)) +
  geom_density() +
  facet_wrap(~ race, nrow = 3)


df %>%
  group_by(parish_name) %>% 
  summarise(n = n()) %>% 
  mutate(parish_name = fct_reorder(parish_name, n)) %>% 
  ggplot(aes(n, parish_name, fill = parish_name)) +
  geom_col(show.legend = F) +
  labs(x = "Number of baptisms in sample",
       y = NULL)

```

### Time on a different axis - shows evolution over time.

- skimr skim
- points
- lines
- arrows
- labels for year with modulo operator
- scale colour manual with colour picker.
- colour of labels with 'colour = "black"'

```{r}
df <- read_rds("data/df_clean_for_export.rds")


df_time <- df %>% 
  filter(country %in% c("Denmark", "Greece"),
         indicator %in% c("Total fertility rate", "Proportion of live births outside marriage")) %>% 
  pivot_wider(names_from = indicator) %>% 
  rename(tfr = `Total fertility rate`, 
         pbom = `Proportion of live births outside marriage`) %>% 
  select(-flag)



df_time %>% 
  mutate(year_flag = ifelse(test = year %% 7 == 0, 
                            yes = year, 
                            no = NA)) %>% 
  ggplot(aes(tfr, pbom, colour = country, label = year_flag)) +
  geom_point() +
  geom_segment(aes(x= tfr, 
                   y= pbom,
                   xend=c(tail(tfr, n=-1), NA), 
                   yend=c(tail(pbom, n=-1), NA)),
               arrow = arrow(length=unit(0.25,"cm"))) +
  geom_text_repel(colour = "black")


```

## gghighlight build up

```{r}
df <- read_rds("data/df_clean_for_export.rds")

df <- df %>% 
  filter(indicator %in% c("Proportion of live births outside marriage")) %>% 
  pivot_wider(names_from = indicator) %>% 
  rename(pbom = `Proportion of live births outside marriage`)

# write_rds(df, "data/births_outside_marriage.rds")

df <- read_rds("data/births_outside_marriage.rds")
  
df %>%   
  filter(flag == 1) %>% 
  ggplot(aes(x = year, y = pbom, group = country, colour = country)) +
  geom_line() +
  scale_y_continuous(labels = percent_format(scale = 1)) +
  labs(y = "Proportion of births\noutside of marriage",
       x = NULL,
       colour = "Country")

```

### Average

```{r}
df %>% 
  group_by(year) %>% 
  mutate(mean_pbom = mean(pbom, na.rm = T)) %>% 
  ungroup() %>% 
  ggplot() +
  geom_line(aes(x = year, y = pbom, group = country, colour = country), 
            data = df %>% filter(flag == 1)) +
  geom_line(aes(x = year, y = mean_pbom, colour = "European average")) +
  scale_y_continuous(labels = percent_format(scale = 1)) +
  labs(y = "Proportion of births\noutside of marriage",
       x = NULL,
       colour = "Country")


```

### Geom ribbon

```{r}
df %>% 
  group_by(year) %>% 
  mutate(mean_pbom = mean(pbom, na.rm = T),
         pct_10 = quantile(pbom, .1, na.rm = T),
         pct_90 = quantile(pbom, .9, na.rm = T)) %>% 
  ungroup() %>% 
  ggplot() +
  geom_line(aes(x = year, y = pbom, group = country, colour = country), 
            data = df %>% filter(flag == 1)) +
  geom_ribbon(aes(x = year, ymin = pct_10, ymax = pct_90, 
                  fill = "10th to 90th\npercentile interval\nfor Europen countries"), alpha = .3) +
  geom_line(aes(x = year, y = mean_pbom, colour = "European average")) +
  scale_y_continuous(labels = percent_format(scale = 1)) +
  scale_fill_manual(values = "#90ee90") +
  guides(fill = guide_legend(order = 2),
         col = guide_legend(order = 1)) +
  labs(y = "Proportion of births\noutside of marriage",
       x = NULL,
       colour = "Country",
       fill = "")

```

### gghighlight

```{r}
library(gghighlight)

df %>%   
  ggplot() +
  geom_line(aes(x = year, y = pbom, colour = country)) +
  gghighlight(flag == 1) +
  scale_y_continuous(labels = percent_format(scale = 1)) +
  theme(legend.text = element_text(size = 4),
        legend.position = "bottom",
        legend.title = element_markdown()) +
  scale_colour_manual(values = c("#C60C30", "#0D5EAF")) +
  labs(title = "Proportion of births outside of marriage in <span style='color:#C60C30'>Denmark</span> and <span style='color:#0D5EAF'>Greece</span>",
       y = "Proportion of births\noutside of marriage",
       x = NULL,
       colour = "Country")

```


### Tie fighter plot

### Heatmap?


## Mapping coffee across the world

### Coffee ratings data prep

```{r}
coffee_ratings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv')

coffee_ratings <- coffee_ratings %>% as_tibble()

coffee_ratings <- coffee_ratings %>% 
  mutate(country_code = countrycode::countrycode(coffee_ratings$country_of_origin, origin = "country.name", destination = "iso3c"),
         iso_a3 = country_code)

coffee_ratings <- coffee_ratings %>% 
  group_by(country_code, species) %>% 
  mutate(mean_country_rating = mean(total_cup_points)) %>% 
  ungroup() %>% 
  distinct(iso_a3, mean_country_rating, species, .keep_all = T)

p_load(rnaturalearth, rnaturalearthdata, rgeos)

world <- ne_countries(scale = "small", returnclass = "sf")

world_df <- inner_join(coffee_ratings, world, by = "iso_a3")

world_df <- world_df %>% 
  select(geometry, iso_a3, mean_country_rating, species, region_un) %>% 
  drop_na()

# write_rds(world_df, "data/coffee_data.rds")
```

## Coffee map

```{r}
ggplot(data = world_df) +
  borders() +
  geom_sf(aes(fill = mean_country_rating, geometry = geometry)) +
  scale_fill_viridis_c(trans = "sqrt") +
  theme_void() +
  theme(legend.position = "bottom") +
  coord_sf(ylim = c(-50, 80)) +
  facet_wrap(~ species, nrow = 2) +
  labs(title = "Average Coffee Rating",
       subtitle = "By country and species",
       fill = "Average Rating / 100",
       caption = "Data: James LeDoux - https://github.com/jldbc/coffee-quality-database",
       x = "",
       y = "")

```

### Coffee columns

```{r}
library(tidytext)

world_df %>% 
  mutate(country_name = countrycode::countrycode(iso_a3, "iso3c", "country.name")) %>%
  mutate(country_name = reorder_within(country_name, mean_country_rating, species)) %>% 
  group_by(species) %>% 
  slice_max(mean_country_rating, n = 10) %>% 
  ungroup() %>% 
  ggplot(aes(x = mean_country_rating, y = country_name, fill = region_un)) +
  geom_col() +
  facet_wrap(~ species, scales = "free") +
  scale_y_reordered() +
  coord_cartesian(xlim = c(70, 90)) +
  labs(x = "Mean coffee rating",
       y = NULL,
       fill = "Region")

```

### map of tiles

```{r}
df <- read.csv("data/worldtilegrid.txt")

df <- df %>% as_tibble() %>% mutate(y = abs(y-23)) 

df <- df %>% 
  rename(iso_a3 = alpha.3) %>% 
  left_join(coffee_ratings %>% 
              filter(species == "Arabica") %>% 
              select(iso_a3, mean_country_rating), by = "iso_a3")

# write_rds(df, "data/coffee_map_tiles.rds")

df %>% 
  ggplot(aes(xmin = x, ymin = y, xmax = x + 1, ymax = y + 1)) +
  geom_rect(color = "#ffffff") +
  geom_rect(color = "#ffffff", aes(fill = mean_country_rating)) +
  scale_fill_viridis_c(trans = "sqrt") +
  geom_text(aes(x = x, y = y, label = alpha.2), color = "#ffffff", 
              alpha = 0.5, nudge_x = 0.5, nudge_y = 0.5, size = 3) +
  theme(panel.grid = element_blank()) +
  theme_void() +
  theme(legend.position = "bottom") +
  coord_equal() +
  labs(fill = "Average Arabica coffee rating by country")

```

```{r}
df %>% 

  ggplot(aes(xmin = x, ymin = y, xmax = x + 1, ymax = y + 1)) +
  geom_rect(color = "#ffffff") +
    geom_text(aes(x = x, y = y, label = alpha.2), color = "#ffffff", 
              alpha = 0.5, nudge_x = 0.5, nudge_y = 0.5, size = 3) +
  theme(panel.grid = element_blank()) +
  theme_void()

```

